\documentclass[a4paper,12pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.6em}

% Assembly style
\lstdefinestyle{AsmStyle}{
    language={[x86masm]Assembler},
    basicstyle=\ttfamily\footnotesize,
    keywordstyle=\color{blue},
    commentstyle=\color{green!40!black},
    numbers=left,
    numberstyle=\tiny,
    breaklines=true,
    frame=single
}

\title{Project 2 Report: \texttt{schedsim}\\
\large CMPE 230 Systems Programming -- Spring 2026}

\author{
Student 1 Name, Student 1 ID \\
Student 2 Name, Student 2 ID
}

\date{\today}

\begin{document}

\maketitle

\section*{Report Guidelines}

This report must explain \textbf{how your assembly program works internally}.

Avoid generic statements. Focus on:
\begin{itemize}
\item how parsing is implemented at register level,
\item how program state is stored in memory,
\item how scheduling decisions are computed,
\item how control flow is structured in assembly.
\end{itemize}


\section{Project Overview}

Briefly describe:
\begin{itemize}
\item what your scheduler does,
\item supported algorithms (FCFS, SJF, SRTF, PF, RR),
\item overall execution flow from input to output.
\end{itemize}


\section{Parsing Logic}

\subsection{Input Processing Strategy}

Explain how you parse the single input line:

\begin{itemize}
\item how tokens are separated (spaces, hyphens),
\item how algorithm type is detected,
\item how numeric values are parsed from characters.
\end{itemize}

\subsection{Register-Level Parsing Design}

Describe explicitly:

\begin{itemize}
\item which registers hold pointers (e.g., input buffer),
\item which registers hold temporary numeric values,
\item how digits are accumulated into integers,
\item how delimiters (' ', '-') are handled.
\end{itemize}

Explain how your parsing avoids corruption and maintains correctness.

\section{State Representation (Memory Layout)}

Explain how all scheduling data is stored.

\subsection{Process Representation}

Describe how each process is represented:

\begin{itemize}
\item ID (character),
\item burst time,
\item arrival time,
\item remaining time,
\item priority (if applicable).
\end{itemize}

Explain whether you used:
\begin{itemize}
\item arrays,
\item structs (manual layout),
\item separate parallel arrays.
\end{itemize}

\subsection{Global State}

Explain:
\begin{itemize}
\item current time,
\item ready queue / process list,
\item output buffer,
\item counters (remaining processes).
\end{itemize}

\subsection{Mandatory Diagram (Hand-Drawn)}

You must include a \textbf{hand-drawn diagram} (scanned or clearly photographed) that shows \textbf{one scheduling step (one clock cycle)} of your program.

Digital diagrams are NOT accepted.

Your diagram must include:

\begin{itemize}
\item memory layout of your process data,
\item the output buffer state,
\item current time and key variables,
\item at least 3-4 registers and what they store,
\item which process is selected in this step,
\item arrows showing how data moves (loads, comparisons, updates).
\end{itemize}

The diagram must correspond to a \textbf{concrete example input}, not an abstract description.

\textbf{Important:}
\begin{itemize}
\item Simply drawing boxes is not sufficient, you must show values.
\item Registers must be labeled with their roles (not just names).
\item The diagram should make it possible to follow your execution step-by-step.
\end{itemize}

\section{Register Usage Strategy}

Clearly document your register design:

\begin{itemize}
\item which registers are used for:
    \begin{itemize}
    \item loop counters,
    \item pointers,
    \item current process index,
    \item temporary computations,
    \end{itemize}
\item which registers are preserved across function calls (if any),
\item how you avoid accidental overwrites.
\end{itemize}

If you used conventions (caller-saved / callee-saved), explain them.

\section{Scheduling Logic}

\subsection{Clock Cycle Simulation}

Explain how one time step is executed:

\begin{itemize}
\item how eligible processes are determined,
\item how the next process is selected,
\item how output is appended,
\item how state is updated.
\end{itemize}

\subsection{Algorithm Implementations}

Briefly explain each:

\begin{itemize}
\item FCFS: arrival-based ordering
\item SJF: sorting or selection by burst
\item SRTF: per-cycle minimum remaining time
\item PF: priority + tie-breaking
\item RR: queue rotation + quantum handling
\end{itemize}

Focus on differences in control flow and data updates.

\section{Control Flow Design}

Describe how your assembly program is structured:

\begin{itemize}
\item main loop (clock cycle loop),
\item parsing phase,
\item scheduling phase,
\item output phase,
\item termination condition.
\end{itemize}

Mention use of:
\begin{itemize}
\item labels,
\item jumps (conditional / unconditional),
\item loop patterns.
\end{itemize}


\section{Representative Code Snippet}

\begin{lstlisting}[style=AsmStyle]
# Example snippet
\end{lstlisting}

Explain:
\begin{itemize}
\item what the code does,
\item which registers are used,
\item how control flow works,
\item why it is correct.
\end{itemize}

\section{Testing and Debugging}

\subsection{Testing Strategy}

\begin{itemize}
\item sample inputs,
\item edge cases (ties, idle CPU, preemption),
\item manual vs automated testing.
\end{itemize}

\subsection{Debugging Approach}

\begin{itemize}
\item print-based debugging,
\item register inspection,
\item step-by-step tracing.
\end{itemize}

\subsection{Edge Cases}

Examples:

\begin{itemize}
\item multiple arrivals at same time,
\item identical priorities,
\item quantum partially used (RR),
\item CPU idle periods.
\end{itemize}

\section{Course Concepts}

Relate your implementation to course topics:

\subsection{Memory and Data Layout}

How memory is organized and accessed.

\subsection{Registers and Low-Level Computation}

How computation is performed without high-level abstractions.

\subsection{Control Flow and Branching}

How loops and decisions are implemented.

\subsection{Parsing Without Libraries}

Manual string processing in assembly.

\subsection{Tradeoffs}

\begin{itemize}
\item simplicity vs efficiency,
\item readability vs performance,
\item static vs dynamic design.
\end{itemize}

\section{Conclusion}

Summarize:

\begin{itemize}
\item strongest design decision,
\item biggest challenge,
\item possible improvements.
\end{itemize}

\section*{AI Usage Statement}

The report must be written by you.

If AI tools were used:
\begin{itemize}
\item which tools,
\item for what purpose,
\item confirmation that all technical content is understood.
\end{itemize}

\newpage

\section*{Appendix: Bad vs Good Report Examples}

This section illustrates the difference between weak and strong reports.

\subsection*{Example 1: Parsing Logic}

\textbf{Bad:}
\begin{quote}
We parse the input string and extract the values. We split by spaces and hyphens and convert numbers.
\end{quote}

\textbf{Why this is bad:}
\begin{itemize}
\item No mention of registers
\item No explanation of how parsing actually works
\item Could be written without implementing anything
\end{itemize}

\textbf{Good:}
\begin{quote}
We parse the input using a pointer stored in \texttt{rsi}, which traverses the input buffer character by character. 
Digits are accumulated into integers using \texttt{rax = rax * 10 + (char - '0')}. 
Hyphens (`-`) are used as field separators and trigger storing the parsed value into the corresponding array. 
Spaces indicate transition to the next process descriptor. 

We ensure correctness by resetting the accumulator register after each field and advancing the pointer carefully to avoid skipping characters.
\end{quote}

\subsection*{Example 2: State Representation}

\textbf{Bad:}
\begin{quote}
We store processes in arrays and update their values during execution.
\end{quote}

\textbf{Why this is bad:}
\begin{itemize}
\item No structure
\item No memory layout
\item No connection to execution
\end{itemize}

\textbf{Good:}
\begin{quote}
Each process is represented using parallel arrays:
\texttt{proc\_id[]}, \texttt{burst[]}, \texttt{arrival[]}, and \texttt{remaining[]}. 

The base address of each array is stored in registers (\texttt{rbx}, \texttt{rcx}, etc.), and process index \texttt{i} is used with scaled indexing to access elements. 

This design simplifies iteration and allows efficient comparisons when selecting the next process. 

\end{quote}

\subsection*{Example 3: Scheduling Logic}

\textbf{Bad:}
\begin{quote}
We implemented SRTF by selecting the process with the smallest remaining time at each step.
\end{quote}

\textbf{Why this is bad:}
\begin{itemize}
\item Just restates the algorithm
\item No implementation detail
\end{itemize}

\textbf{Good:}
\begin{quote}
At each clock cycle, we iterate over all processes and compare their remaining times. 
The current minimum is stored in \texttt{r8}, and the index of the selected process is stored in \texttt{r9}. 

For each process:
\begin{itemize}
\item we check if it has arrived,
\item we skip completed processes (remaining = 0),
\item we update the minimum if a smaller remaining time is found.
\end{itemize}

After selection, the chosen process's remaining time is decremented and its ID is written into the output buffer. 

\end{quote}

\subsection*{Example 4: Register Usage}

\textbf{Bad:}
\begin{quote}
We used registers to store temporary values.
\end{quote}

\textbf{Why this is bad:}
\begin{itemize}
\item meaningless
\item no mapping
\end{itemize}

\textbf{Good:}
\begin{quote}
We use:
\begin{itemize}
\item \texttt{rsi} as input pointer,
\item \texttt{rdi} as output buffer pointer,
\item \texttt{rcx} as loop counter,
\item \texttt{r8} for current minimum remaining time,
\item \texttt{r9} for selected process index.
\end{itemize}

Registers are reused carefully, and critical values are preserved across loops to prevent corruption.
\end{quote}

\subsection*{Summary}

Strong reports:
\begin{itemize}
\item explain \textbf{how}, not just \textbf{what}
\item refer to registers, memory, and control flow
\item include concrete details and examples
\end{itemize}

Weak reports:
\begin{itemize}
\item restate the assignment
\item avoid implementation details
\item could be written without coding the project
\end{itemize}
\end{document}
